---
title: "Logistic Regression"
author: "Emmanuel Kasigazi"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: console
  markdown: 
    wrap: 72
---

This lab introduces simple logistic regression, a model for the
association of a binary response variable with a single predictor
variable. Logistic regression generalizes methods for two-way tables and
allows for the use of a numerical predictor.

The material in this lab corresponds to Section 9.xx in *OpenIntro
Statistics*.

### Introduction

*Odds and probabilities*

If the probability of an event $A$ is $p$, the odds of the event are
$$\dfrac{p}{1 - p}. $$

Given the odds of an event $A$, the probability of the event is
$$\dfrac{\text{odds}}{1 + \text{odds}}. $$

*Simple logistic regression*

Suppose that $Y$ is a binary response variable and $X$ is a predictor
variable, where $Y = 1$ represents the particular outcome of interest.

The model for a single variable logistic regression, where
$p(x) = P(Y = 1 | X = x)$, is
$$\text{log} \left[ \dfrac{p(x)}{1 - p(x)} \right] = \beta_0 + \beta_1 x. $$

The estimated model equation has the form
$$\text{log} \left[ \dfrac{\hat{p}(x)}{1 - \hat{p}(x)} \right] = b_0 + b_1 x, $$
where $b_0$ and $b_1$ are estimates of the model parameters $\beta_0$
and $\beta_1$.

### Background Information

Patients admitted to an intensive care unit (ICU) are either extremely
ill or considered to be at great risk of serious complications. There
are no widely accepted criteria for distinguishing between patients who
should be admitted to an ICU and those for whom admission to other
hospital units would be more appropriate. Thus, among different ICUs,
there are wide ranges in a patient's chance of survival. When studies
are done to compare effectiveness of ICU care, it is critical to have a
reliable means of assessing the comparability of the different patient
populations.

One such strategy for doing so involves the use of statistical modeling
to relate empirical data for many patient variables to outcomes of
interest. The following dataset consists of a sample of 200 subjects who
were part of a much larger study on survival of patients following
admission to an adult
ICU.\footnote{From Hosmer D.W., Lemeshow, S., and Sturdivant, R.X. \textit{Applied Logistic Regression}. 3$^{rd}$ ed., 2013.}
The major goal of the study was to develop a logistic regression model
to predict the probability of survival to hospital
discharge.\footnote{Lemeshow S., et al. Predicting the outcome of intensive care unit patients. \textit{Journal of the American Statistical Association} 83.402 (1988): 348-356.}

The following table provides a list of the variables in the dataset and
their description. The data are accessible as the \texttt{icu} dataset
in the \texttt{aplore3} package.

\begin{center}
\begin{tabular}{r|l}
\textbf{Variable} & \textbf{Description} \\
\hline
\texttt{id} & patient ID number \\
\texttt{sta} & patient status at discharge, either \texttt{Lived} or \texttt{Died} \\
\texttt{age} & age in years (when admitted) \\
\texttt{gender} & gender, either \texttt{Male} or \texttt{Female} \\
\texttt{can} & cancer part of current issue, either \texttt{No} or \texttt{Yes} \\
\texttt{crn} & history of chronic renal failure, either \texttt{No} or \texttt{Yes}\\
\texttt{inf} & infection probable at admission, either \texttt{No} or \texttt{Yes} \\
\texttt{cpr} & CPR prior to admission, either \texttt{No} or \texttt{Yes} \\
\texttt{sys} & systolic blood pressure at admission, in mm Hg \\
\texttt{hra} & heart rate at admission, in beats per minute \\
\texttt{pre} & previous admission to an ICU within 6 months, either \texttt{No} or \texttt{Yes} \\
\texttt{type} & type of admission, either \texttt{Elective} or \texttt{Emergency} \\
\texttt{fra} & long bone, multiple, neck, single area, or hip fracture, either \texttt{No} or \texttt{Yes} \\
\texttt{po2} & $PO_2$ from initial blood gases, either \texttt{60} or \texttt{<=60}, in mm Hg\\
\texttt{ph} & $pH$ from initial blood gases, either \texttt{>= 7.25} or \texttt{< 7.25} \\
\texttt{pco} & $PCO_2$ from initial blood gases, either \texttt{<= 45} or \texttt{>45}, in mm Hg \\
\texttt{bic} & $HCO_3$ (bicarbonate) from initial blood gases, either \texttt{>= 18} or \texttt{<18}, in mm Hg \\
\texttt{cre} & creatinine from initial blood gases, either \texttt{<= 2.0} or \texttt{> 2.0}, in mg/dL \\
\texttt{loc} & level of consciousness at admission, either \texttt{Nothing}, \texttt{Stupor}, or \texttt{Coma}
\end{tabular}
\end{center}

### Odds and probabilities

1.  Create a two-way table of survival to discharge by whether CPR was
    administered prior to admission. The template provides code for
    re-leveling the \texttt{sta} variable such that \texttt{0}
    corresponds to \texttt{Died} and \texttt{1} corresponds to
    \texttt{Lived}.

```{r, eval = FALSE}
#install the package (only needs to be done once)
install.packages("aplore3")
```

```{r, warning = FALSE}
#load the data
library(aplore3)
data("icu")

#relevel survival so that 1 corresponds to surviving to discharge
icu$sta = factor(icu$sta, levels = rev(levels(icu$sta)))

#create two-way table
addmargins(table(icu$sta, icu$cpr,
                 dnn = c("Survival", "Prior CPR")))
```

a)  Calculate the odds of survival to discharge for those who did not
    receive CPR prior to ICU admission. Is someone who did not receive
    CPR prior to admission more likely to survive to discharge than to
    not survive to discharge?

```{r}
#use r as a calculator
# Using the table that was already created
cpr_table <- table(icu$sta, icu$cpr)

# Calculate odds of survival for patients without prior CPR
odds_no_cpr <- cpr_table["Lived", "No"] / cpr_table["Died", "No"]

# Print the odds
odds_no_cpr

# Check if odds are greater than 1
odds_no_cpr > 1
```

```         
The odds of survival for patients without prior CPR are:
```

Odds = Number who survived / Number who died Odds = 154 / 33 Odds = 4.67

This means that for patients who did not receive CPR prior to ICU
admission, the odds of survival to discharge are 4.67 to 1.

Since the odds are greater than 1 (4.67 \> 1), patients who did not
receive CPR prior to admission are more likely to survive to discharge
than to not survive to discharge. In fact, they are about 4.67 times
more likely to survive than to die.

b)  Calculate the odds of survival to discharge for those who received
    CPR prior to ICU admission. Is someone who received CPR prior to
    admission more likely to survive to discharge than not?

```{r}
#use r as a calculator
# Using the table that was already created
cpr_table <- table(icu$sta, icu$cpr)

# Calculate odds of survival for patients with prior CPR
odds_yes_cpr <- cpr_table["Lived", "Yes"] / cpr_table["Died", "Yes"]

# Print the odds
odds_yes_cpr

# Check if odds are greater than 1
odds_yes_cpr > 1

```

Since these odds are less than 1 (0.857 \< 1), patients who received CPR
prior to ICU admission are NOT more likely to survive to discharge than
to die.

The odds ratio of 0.857 means that for every 10 patients with prior CPR
who die, approximately 8-9 patients with prior CPR survive.

c)  Calculate the odds ratio of survival to discharge, comparing
    patients who receive CPR prior to admission to those who do not
    receive CPR prior to admission.

```{r}
#use r as a calculator
    
# Using the table that was already created
cpr_table <- table(icu$sta, icu$cpr)

# Calculate odds of survival for each group
odds_yes_cpr <- cpr_table["Lived", "Yes"] / cpr_table["Died", "Yes"]
odds_no_cpr <- cpr_table["Lived", "No"] / cpr_table["Died", "No"]

# Calculate odds ratio (comparing CPR Yes to CPR No)
odds_ratio <- odds_yes_cpr / odds_no_cpr

# Print the odds ratio
odds_ratio

```

The odds ratio is: OR = 0.857/4.667 = 0.184 This means that the odds of
survival for patients who received CPR prior to admission are only about
18.4% of the odds for patients who didn't receive CPR.

In other words, patients who received CPR prior to admission have
approximately 5.4 times lower odds of survival compared to patients who
didn't receive CPR.

2.  Creatinine level in the data are recorded as being either less than
    or equal to 2.0 mg/dL or greater than 2.0 mg/dL. A typical
    creatinine level is between 0.5 - 1.0 mg/dL, and elevated creatinine
    may be a sign of renal failure.

```{r}
#create two-way table  table for creatinine and survival

addmargins(table(icu$sta, icu$cre, 
                 dnn = c("Survival", "Creatinine")))
```

```{r}
#Odds of survival for patients with creatinine ≤2.0 mg/dL
odds_normal_cre <- 155/35
odds_normal_cre  # This equals 4.43
```

```{r}
#Odds of survival for patients with creatinine >2.0 mg/dL
odds_high_cre <- 5/5
odds_high_cre  # This equals 1.0
```

```{r}
#Odds ratio comparing high creatinine to normal creatinine:
odds_ratio <- odds_high_cre/odds_normal_cre
odds_ratio  # This equals approximately 0.226
```

This means:

-Patients with normal creatinine (≤2.0 mg/dL) have 4.43 times higher
odds of surviving than dying

-Patients with elevated creatinine (\>2.0 mg/dL) have equal odds of
surviving or dying (1:1) -The odds of survival for patients with
elevated creatinine are only about 22.6% of the odds for patients with
normal creatinine

-Only 5% of patients in the sample (10/200) had elevated creatinine
levels

These results suggest that elevated creatinine, which can indicate
kidney dysfunction, is associated with reduced odds of survival in ICU
patients

a)  Calculate the odds of survival to discharge for patients who have a
    creatinine level less than or equal to 2.0 mg/dL. From the odds,
    calculate the probability of survival to discharge for these
    patients.

```{r}
#use r as a calculator
# From the table data
survived_normal_cre <- 155
died_normal_cre <- 35

# Calculate odds:  Odds of survival for patients with creatinine ≤2.0 mg/dL:
odds_normal_cre <- survived_normal_cre / died_normal_cre
odds_normal_cre  # This equals 4.43
```

The odds of survival for patients with creatinine ≤2.0 mg/dL are 4.43 to
1.

```{r}
# Convert odds to probability
probability_normal_cre <- odds_normal_cre / (1 + odds_normal_cre)
probability_normal_cre  # This equals approximately 0.816
```

The probability of survival to discharge for patients with creatinine
≤2.0 mg/dL is approximately 0.816 or 81.6%.

```{r}
#This can also be calculated directly from the table as the proportion:
probability_normal_cre <- 155 / 190
probability_normal_cre  # This equals approximately 0.816
```

Patients admitted to ICU with creatinine levels of ≤ 2.0 mg/dL have a
high probability (81.6%) of surviving to discharge, reflecting
relatively less severe renal impairment at admission.

b)  Calculate the probability of survival to discharge for patients who
    have a creatinine level greater than 2.0 mg/dL. From the
    probability, calculate the odds of survival to discharge for these
    patients.

```{r}
#use r as a calculator
#calculate the probability directly from the table:
# From the table data
survived_high_cre <- 5
total_high_cre <- 10

# Calculate probability of survival
probability_high_cre <- survived_high_cre / total_high_cre
probability_high_cre  # This equals 0.5 or 50%

```

```{r}
# Convert probability to odds
odds_high_cre <- probability_high_cre / (1 - probability_high_cre)
odds_high_cre  # This equals 1.0
```

Patients with creatinine levels greater than 2.0 mg/dL have a
significantly reduced probability (50%) of survival compared to those
with lower creatinine levels.

Their odds of survival (1.0) reflect equal likelihood of survival and
non-survival, signaling substantially higher clinical risk or illness
severity.

c)  Compute and interpret the odds ratio of survival to discharge,
    comparing patients with creatinine $> 2.0$ mg/dL to those with
    creatinine $\leq$ 2.0 mg/dL.

```{r}
#use r as a calculator
# Calculate odds ratio (comparing high creatinine to normal creatinine)
odds_ratio <- 1.0 / 4.43
odds_ratio  # This equals approximately 0.226
```

Odds Ratio = 0.226 Interpretation: The odds ratio of 0.226 means that
patients with creatinine levels \>2.0 mg/dL have approximately 22.6% of
the odds of survival compared to patients with creatinine levels ≤2.0
mg/dL.

In other words, patients with high creatinine levels have approximately
4.43 times lower odds of survival compared to those with normal or
moderately elevated creatinine.

This suggests that severely elevated creatinine levels (\>2.0 mg/dL),
which can indicate significant kidney dysfunction, are associated with
substantially reduced odds of survival in ICU patients.

This makes clinical sense, as renal failure is a serious condition that
can complicate critical illness and increase mortality risk.

### Simple logistic regression

3.  Fit a logistic regression model to predict survival to discharge
    from prior CPR.

```{r}
#fit a model
glm(sta ~ cpr, data = icu, family = binomial(link = "logit"))
```

a)  Write the model equation estimated from the data.

The estimated model equation from the data is: log[p̂(x)/(1-p̂(x))] =
1.540 - 1.695 × cprYes

Where:

a)  p̂(x) represents the estimated probability of survival to discharge
b)  cprYes is 1 if the patient received CPR prior to admission and 0 if
    they did not
c)  1.540 is the intercept (β₀)
d)  -1.695 is the coefficient for prior CPR (β₁)

This logistic regression equation models the log odds of survival based
on whether the patient received CPR prior to ICU admission.

b)  Interpret the intercept. Confirm that the interpretation coheres
    with the answer to Question 1, part a).

The intercept (β₀) in our model is 1.540. Interpretation of the
intercept:

The intercept represents the log odds of survival to discharge for
patients in the reference group - in this case, patients who did NOT
receive CPR prior to admission (when cprYes = 0). To interpret this
value:

Log odds of survival for no CPR = 1.540 Converting from log odds to
odds: e\^1.540 = 4.67

This means that for patients who did not receive CPR prior to admission,
the odds of survival to discharge are 4.67 to 1. This interpretation
coheres perfectly with our answer to Question 1, part a), where we
calculated the odds of survival for patients without prior CPR as: 154
(survived) / 33 (died) = 4.67

Both calculations give us the exact same odds value, confirming that the
intercept in the logistic regression model correctly represents the log
odds of survival for patients who did not receive CPR before admission.

Confirming with Question 1 (part a): Previously, in Question 1 part (a),
we calculated the odds and probability of survival for patients without
prior CPR directly from the two-way table:

Odds previously calculated: 154/33≈4.67 154/33≈4.67

Probability previously calculated: about 82.4%.

This aligns very closely with the logistic regression intercept results
(Odds ≈ 4.66, Probability ≈ 82.3%).

The logistic regression intercept interpretation perfectly coheres with
the earlier direct calculation from Question 1(a).

Both analyses consistently indicate a survival probability around 82%
for patients who did not receive CPR prior to ICU admission.

c)  Interpret the slope coefficient. Compute the exponential of the
    slope coefficient and confirm that this matches the answer to
    Question 1, part c).

The slope coefficient (β₁) in our model is -1.695. Interpretation of the
slope coefficient:

The slope coefficient represents the change in log odds of survival when
comparing patients who received CPR prior to admission (cprYes = 1) to
those who did not (cprYes = 0). To interpret this value:

-Change in log odds = -1.695

-Converting to an odds ratio: e\^(-1.695) = 0.184

This means that the odds of survival for patients who received CPR prior
to admission are 0.184 times the odds for patients who did not receive
CPR. In other words, patients who received CPR have approximately 81.6%
lower odds of survival compared to those who didn't.

Computing the exponential of the slope coefficient: e\^(-1.695) = 0.184

This matches exactly with our answer to Question 1, part c), where we
calculated the odds ratio as: (6/7) ÷ (154/33) = 0.857 ÷ 4.667 = 0.184

Both calculations give us the same odds ratio, confirming that the slope
coefficient in the logistic regression model correctly represents the
log odds ratio comparing survival between patients who received CPR and
those who didn't.

d)  Compute and interpret an odds ratio that summarizes the association
    between survival to discharge and prior CPR.

```{r}
#compute odds ratio
# Using the contingency table approach
cpr_table <- table(icu$sta, icu$cpr)
odds_ratio <- (cpr_table["Lived", "Yes"]/cpr_table["Died", "Yes"]) / 
              (cpr_table["Lived", "No"]/cpr_table["Died", "No"])
odds_ratio
```

The odds ratio (0.184) means patients who received CPR prior to ICU
admission have about 81.6% lower odds of survival compared to patients
who did not receive prior CPR.

This clearly indicates a strong negative association, suggesting that
prior CPR reflects severe illness or critical condition associated with
lower survival likelihood.

e)  Is there evidence of a statistically significant association between
    survival to discharge and prior CPR at $\alpha = 0.05$?

```{r}
#use summary(glm( ))
# Fit the model
cpr_model <- glm(sta ~ cpr, data = icu, family = binomial)

# Get the summary with p-values
summary(cpr_model)
```

The p-value for the cprYes coefficient is 0.00398, which is less than
the significance level of α = 0.05. Therefore, there is evidence of a
statistically significant association between survival to discharge and
prior CPR at α = 0.05.

We can reject the null hypothesis that there is no association between
prior CPR and survival.

The output shows:

-cprYes coefficient = -1.6946

-Standard error = 0.5885

-z-value = -2.880

-p-value = 0.00398

This significant negative coefficient confirms that patients who
received CPR prior to ICU admission have significantly lower odds of
survival compared to those who did not receive CPR.

4.  Fit a logistic regression model to predict survival to discharge
    from an indicator of elevated creatinine.

```{r}
#fit the model
# Fit logistic regression model with creatinine as predictor
cre_model <- glm(sta ~ cre, data = icu, family = binomial(link = "logit"))

# Display model summary
summary(cre_model)
```

a)  Write the model equation estimated from the data.

The model equation is: log[p̂(x)/(1-p̂(x))] = 1.4881 - 1.4881 × cre\>2.0

Where:

-The intercept (1.4881) represents the log odds of survival for patients
with creatinine ≤2.0 mg/dL

-The coefficient for elevated creatinine (-1.4881) indicates that having
creatinine \>2.0 mg/dL is associated with lower odds of survival

-The p-value for the creatinine coefficient (0.0241) is statistically
significant at α=0.05

Converting to odds:

-For normal creatinine: e\^1.4881 = 4.43 (matching our earlier
calculation of 155/35)

-For elevated creatinine: e\^(1.4881-1.4881) = e\^0 = 1.0 (matching our
earlier calculation of 5/5)

The odds ratio comparing elevated to normal creatinine is: e\^(-1.4881)
= 0.226

This means patients with elevated creatinine have approximately 22.6% of
the odds of survival compared to those with normal creatinine levels.

b)  Interpret the intercept and slope coefficient.

Interpreting the intercept and slope coefficient from the logistic
regression model:

Intercept (1.4881): The intercept represents the log odds of survival to
discharge for patients in the reference group - those with creatinine
≤2.0 mg/dL.

Converting this to odds: e\^1.4881 = 4.43, meaning that patients with
normal or moderately elevated creatinine have odds of 4.43 to 1 of
surviving to discharge.

This tells us that these patients are more than four times as likely to
survive than to die.

Slope coefficient (-1.4881): The coefficient for "cre\> 2.0" represents
the change in log odds of survival when comparing patients with
creatinine \>2.0 mg/dL to the reference group.

The negative value indicates that elevated creatinine is associated with
lower odds of survival. Specifically, the log odds of survival decrease
by 1.4881 when a patient has creatinine \>2.0 mg/dL.

Converting this to an odds ratio: e\^(-1.4881) = 0.226, meaning that
patients with elevated creatinine have only 22.6% of the odds of
survival compared to patients with creatinine ≤2.0 mg/dL.

c)  Compute and interpret an odds ratio that summarizes the association
    between survival to discharge and an indicator of elevated
    creatinine.

```{r}
#compute odds ratio
# Compute odds ratio from model coefficient
exp(coef(cre_model)["cre> 2.0"])
```

Interpretation of the odds ratio: The odds ratio of 0.226 means that
patients with elevated creatinine (\>2.0 mg/dL) have approximately 22.6%
of the odds of survival to discharge compared to patients with normal or
moderately elevated creatinine (≤2.0 mg/dL).

In other words, patients with high creatinine levels have approximately
4.43 times lower odds of survival. This odds ratio quantifies the
negative association between elevated creatinine and survival,
suggesting that kidney dysfunction (as indicated by high creatinine) is
associated with substantially reduced odds of survival in ICU patients.

The statistically significant p-value (0.0241) indicates that this
association is unlikely to be due to chance.

In short: Elevated creatinine (\>2.0 mg/dL) is strongly associated with
decreased odds of survival, underscoring its significance as an
indicator of higher mortality risk among ICU patients.

d)  Is there evidence of a statistically significant association between
    survival to discharge and an indicator of elevated creatinine at
    $\alpha = 0.05$?

```{r}
# Fit the model
cre_model <- glm(sta ~ cre, data = icu, family = binomial)

# Extract the p-value for the creatinine coefficient
summary(cre_model)$coefficients["cre> 2.0", "Pr(>|z|)"]

# Compare to alpha level
summary(cre_model)$coefficients["cre> 2.0", "Pr(>|z|)"] < 0.05

```

The p-value for the "cre\> 2.0" coefficient is 0.0241, which is less
than our significance level of α = 0.05.

Therefore, there is evidence of a statistically significant association
between survival to discharge and elevated creatinine at α = 0.05.

We can reject the null hypothesis that there is no association between
elevated creatinine and survival.

This confirms that elevated creatinine levels are significantly
associated with lower odds of survival in ICU patients, which aligns
with clinical expectations since impaired kidney function (as indicated
by high creatinine) often complicates critical illness.

5.  Fit a logistic regression model to predict survival to discharge
    from age.

```{r}
#fit the model
# Fit logistic regression model with age as predictor
age_model <- glm(sta ~ age, data = icu, family = binomial(link = "logit"))

# Display model summary
summary(age_model)

```

Older age is statistically significantly associated with decreased odds
of ICU patient survival.

Age is a meaningful predictor of survival in ICU patients, reflecting
clinical expectations that older patients have generally worse outcomes.

a)  Write the model equation estimated from the data. The fitted
    logistic regression model for predicting survival to discharge from
    age is: log[p̂(x)/(1-p̂(x))] = 3.05851 - 0.02754 × age

This model has:

Intercept (β₀): 3.05851 Age coefficient (β₁): -0.02754

The model shows that age has a statistically significant negative
association with survival (p=0.00913), indicating that older patients
have lower odds of survival to discharge.

For each one-year increase in age, the log odds of survival decrease by
0.02754, which corresponds to a multiplicative change in odds of
e\^(-0.02754) = 0.973.

This means the odds of survival decrease by approximately 2.7% for each
additional year of age. The intercept (3.05851) represents the log odds
of survival for a theoretical patient with age=0, which isn't directly
interpretable in this context since we don't have newborns in an adult
ICU.

b)  Does the intercept have a meaningful interpretation in the context
    of the data?

No, the intercept in this logistic regression model doesn't have a
meaningful interpretation in the context of the data.

The intercept (3.05851) represents the log odds of survival for a
patient with age = 0, which isn't meaningful since:

-The dataset contains adult ICU patients, not newborns

-A patient with age = 0 is outside the range of the observed data

-Extrapolating to age = 0 isn't clinically relevant

When working with continuous predictors like age, the intercept often
lacks practical interpretation because it represents a scenario outside
the realistic domain of the data. Instead, it's more useful to calculate
the predicted log odds or probability of survival for specific,
clinically relevant ages within the range of the observed data.

A more meaningful approach would be to center the age variable around
its mean or a clinically relevant value, which would give the intercept
an interpretable meaning as the log odds for a patient of that reference
age.

c)  Interpret the slope coefficient.

The slope coefficient for age in the model is -0.02754. Interpretation
of the slope coefficient: For each one-year increase in a patient's age,
the log odds of survival to discharge decrease by 0.02754, holding all
other variables constant.

To make this more intuitive, we can convert from log odds to an odds
ratio:

Odds ratio = e\^(-0.02754) = 0.973

This means that for each additional year of age, the odds of survival to
discharge multiply by a factor of 0.973, which represents a decrease of
approximately 2.7% in the odds of survival per year of age.

In clinical terms, this indicates that older ICU patients have
progressively lower odds of survival compared to younger patients, with
each year of age associated with a small but statistically significant
reduction in survival odds. This finding aligns with clinical
expectations, as advanced age is a known risk factor for poorer outcomes
in critical illness.

d)  Calculate the odds of survival to discharge for a 70-year-old
    individual. Is a 70-year-old individual more likely to survive than
    not?

```{r}
#use r as a calculator
# Calculate log odds for a 70-year-old
log_odds_70 <- 3.05851 - 0.02754 * 70

# Convert log odds to odds
odds_70 <- exp(log_odds_70)

# Print the odds
odds_70

# Check if odds are greater than 1
odds_70 > 1

#alternatively, use predict()...


```

This means that a 70-year-old patient has odds of approximately 3.098 to
1 of surviving to discharge. Since these odds are greater than 1 (3.098
\> 1), a 70-year-old individual is more likely to survive than not.

Specifically, for every 70-year-old patient who dies, we expect about
3.098 patients of the same age to survive.

I could also convert this to a probability: Probability = odds/(1+odds)
= 3.098/(1+3.098) = 0.756 This means a 70-year-old patient has
approximately a 75.6% probability of surviving to discharge.Retry

e)  Calculate the odds ratio of survival to discharge comparing a
    45-year-old individual to a 70-year-old individual.

```{r}
#use r as a calculator
# Odds ratio using the slope coefficient directly
age_difference <- 45 - 70
odds_ratio <- exp(-0.02754 * age_difference)
odds_ratio

```

```{r}
#alternatively, use predict()...
# Calculate the log odds for a 45-year-old
log_odds_45 <- 3.05851 - 0.02754 * 45

# Calculate the log odds for a 70-year-old
log_odds_70 <- 3.05851 - 0.02754 * 70

# Calculate odds for each age
odds_45 <- exp(log_odds_45)
odds_70 <- exp(log_odds_70)

# Calculate the odds ratio
odds_ratio_45_vs_70 <- odds_45 / odds_70

# Print the odds ratio
odds_ratio_45_vs_70

```

This odds ratio of 1.99 means that a 45-year-old patient has
approximately 1.99 times the odds of survival to discharge compared to a
70-year-old patient. In other words, the odds of survival for a
45-year-old are about twice those of a 70-year-old.
